---
title: "Splenic erythroid composition across CASP8 genotypes — propeller compositional analysis"
subtitle: "Standalone, fully reproducible script for the publication figure (per-stage fraction of total splenocytes)"
author:
  - name: "Dr. Ali T. Abdallah"
    email: "ali.abdallah@uni-koeln.de"
    affiliation: "CECAD Research Center - Bioinformatics Facility"
date: "`r Sys.Date()`"
version: "1.0"
output:
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,   # show every step so the analysis is reproducible
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)

# Fixed seed for fully reproducible jitter/label placement.
set.seed(42)

# ── Dependencies ────────────────────────────────────────────────────────────
# CRAN:        Seurat, ggplot2, dplyr, tidyr, tibble, ggrepel, scales, qs, knitr
# Bioconductor: speckle, limma
# All steps below run on a single annotated Seurat object and require no
# project-specific helper functions, so this file is self-contained.
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(ggrepel)
library(scales)
library(DT)
library(openxlsx)
```

# Purpose

This report is a **self-contained, publicly shareable** script that reproduces
the barplot used in the manuscript. Significance was assessed by the 
**propeller** compositional test (Phipson *et al.*, *Bioinformatics* 2022).

In this script views of the same propeller analysis are produced:

1. A **per-stage bar plot** of the genotype-level fractions (mean +/- SEM,
   one dot per mouse, propeller p-values), which shows the direction and
   magnitude of the compositional shift. **This is the figure selected for
   the manuscript.**
2. A **QQ plot of propeller t-statistics**, faceted by pairwise contrast,
   retained here as an diagnostic of the per-stage t-statistics.

Everything needed is below, in order:


1. **Input**: load the annotated Seurat object after downloading it from FigShare (See README).
2. **Erythroid stage assignment**: define the three maturation stages.
3. **Compositional test**: the propeller statistics.
4. **Publication figure**: the per-stage bar plot (primary), exported at
   publication quality. A QQ plot of propeller t-statistics is also produced
   as an internal diagnostic.
5. **Statistics table** and **session info**.

---

# Input - annotated Seurat object

```{r load_object}
# Annotated, cleansed Seurat object (all splenocytes, all genotypes).
# Override via the SEURAT_PATH env var (e.g. to point at a reproduced object).
SEURAT_PATH <- Sys.getenv("SEURAT_PATH",
                          unset = "data/seurat_objects.combined.cleansed.annotated.250428.qs")

# Output directory for figures/tables. Override via the FIG_DIR env var (e.g.
# to keep a reproduced run's outputs alongside the reproduced object instead
# of the paper's own figures/ folder).
FIG_DIR <- Sys.getenv("FIG_DIR", unset = "figures")

so <- qs::qread(SEURAT_PATH)

# Per-sample replicate identifier used for genotype assignment and the
# propeller design matrix. Some reproduced objects lose per-sample identity in
# orig.ident (e.g. reset to a single project name during integration); point
# SAMPLE_COL at the metadata column that actually holds it in that case.
SAMPLE_COL <- Sys.getenv("SAMPLE_COL", unset = "orig.ident")
so$orig.ident <- so[[SAMPLE_COL]][[1]]

# Genotype from the sample identifier.
so$genotype <- dplyr::case_when(
  grepl("^WT", so$orig.ident) ~ "WT",
  grepl("^KO", so$orig.ident) ~ "KO",
  grepl("^CS", so$orig.ident) ~ "CS"
)
geno_levels <- c("WT", "KO", "CS")
so$genotype <- factor(so$genotype, levels = geno_levels)

# Fixed colour-blind-considerate genotype palette used throughout the project.
geno_cols <- c(WT = "#4575B4", KO = "#FDAE61", CS = "#D73027")
```

---

# Erythroid stage assignment

The erythroid population is split into three maturation stages by
marker-panel module scoring.

```{r stage_assignment}
# 1. Subset to erythroid cells by cell-type label.
erythroid_pattern   <- "(?i)eryth|erythroblast|proery|reticulocyte"
erythroid_celltypes <- grep(erythroid_pattern, unique(so$cell_type),
                            value = TRUE, perl = TRUE)
so_ery <- subset(so, cell_type %in% erythroid_celltypes)

# 2. Curated marker panels for the three maturation stages.
marker_panels <- list(
  Erythroid_Progenitor = c("Kit","Cd34","Gata2","Ly6a","Myb","Tal1","Lmo2","Lyl1",
                            "Gata1","Epor","Itga2b","Cited2","Meis1"),
  Erythroblast         = c("Gata1","Klf1","Gypa","Car2","Alas2","Tfrc","Nfe2",
                            "Slc25a37","Slc25a38","Fech","Hbb-bh1","Hbb-y"),
  Late_Erythroblast    = c("Hbb-bs","Hba-a1","Hba-a2","Slc4a1","Epb42","Hemgn",
                            "Spta1","Ank1","Bnip3l","Tmcc2","Bpgm","Rhd","Rhag","Add2","Sptb")
)

# Keep only genes present and reliably detected (>= 5% of erythroid cells).
marker_panels <- lapply(marker_panels, function(g) intersect(g, rownames(so_ery)))
marker_panels <- marker_panels[lengths(marker_panels) > 0]

DETECT_MIN  <- 0.05
.detect_rate <- rowMeans(
  GetAssayData(so_ery, assay = "RNA", layer = "counts")[
    unique(unlist(marker_panels)), , drop = FALSE] > 0)
marker_panels <- lapply(marker_panels, function(g) g[.detect_rate[g] >= DETECT_MIN])
marker_panels <- marker_panels[lengths(marker_panels) > 0]

# 3. Module-score each panel and assign each cell to its winning (argmax) panel.
old_cols <- grep("^score_", colnames(so_ery@meta.data), value = TRUE)
if (length(old_cols)) so_ery@meta.data[old_cols] <- NULL
for (stage in names(marker_panels)) {
  so_ery <- AddModuleScore(so_ery,
                           features = list(marker_panels[[stage]]),
                           name     = paste0("score_", stage, "_"))
}
score_cols          <- paste0("score_", names(marker_panels), "_1")
score_mat           <- as.matrix(so_ery@meta.data[, score_cols])
colnames(score_mat) <- names(marker_panels)
winner              <- colnames(score_mat)[apply(score_mat, 1, which.max)]

# 4. Drop low-depth cells (< 2 detected markers from their winning panel).
expr_mat   <- GetAssayData(so_ery, assay = "RNA", layer = "counts")
n_detected <- vapply(seq_along(winner), function(i) {
  panel <- intersect(marker_panels[[winner[i]]], rownames(expr_mat))
  if (!length(panel)) return(0L)
  sum(expr_mat[panel, i] > 0)
}, integer(1))

so_ery$ery_stage <- factor(
  ifelse(n_detected >= 2L, winner, "Unassigned_lowDepth"),
  levels = c(names(marker_panels), "Unassigned_lowDepth")
)
so_ery <- subset(so_ery, ery_stage != "Unassigned_lowDepth")

# Marker panels actually used (post detection-rate filter).
knitr::kable(
  tibble::tibble(
    Stage     = c("Progenitor", "Erythroblast", "Late Erythroblast")[
                   match(names(marker_panels),
                         c("Erythroid_Progenitor","Erythroblast","Late_Erythroblast"))],
    `# genes` = lengths(marker_panels),
    Markers   = vapply(marker_panels, paste, character(1), collapse = ", ")
  ),
  caption = "Stage marker panels used for assignment (after detection-rate filter).",
  align   = c("l", "r", "l")
)
```
---

# Compositional test (propeller)

Proportions are computed relative to **all captured splenocytes**: every
non-erythroid cell is pooled into a single `Other` cluster so that the three
erythroid-stage proportions are fractions of the whole spleen. propeller-test is 
then run on the three pairwise genotype contrasts.

```{r propeller_test}
# Cell-level label spanning ALL splenocytes: stage label for erythroid cells,
# "Other" for everything else.
all_cluster_label <- rep("Other", ncol(so))
names(all_cluster_label) <- colnames(so)
ery_keep <- so_ery@meta.data %>%
  tibble::rownames_to_column("cell") %>%
  dplyr::filter(ery_stage != "Unassigned_lowDepth")
all_cluster_label[ery_keep$cell] <- as.character(ery_keep$ery_stage)

# Per-replicate proportions with arcsine-sqrt variance-stabilising transform.
prop_obj <- speckle::getTransformedProps(
  clusters  = all_cluster_label,
  sample    = so$orig.ident,
  transform = "asin"
)

# Design matrix (one column per genotype) aligned to the proportion columns.
sample_geno_lut <- unique(so@meta.data[, c("orig.ident", "genotype")])
sample_geno_lut$genotype <- factor(sample_geno_lut$genotype, levels = geno_levels)
sample_geno_lut <- sample_geno_lut[match(colnames(prop_obj$TransformedProps),
                                         sample_geno_lut$orig.ident), ]

design_mat <- model.matrix(~ 0 + sample_geno_lut$genotype)
colnames(design_mat) <- levels(sample_geno_lut$genotype)

# Three pairwise contrasts.
contr_mat <- limma::makeContrasts(
  CS_vs_WT = CS - WT,
  CS_vs_KO = CS - KO,
  WT_vs_KO = WT - KO,
  levels   = design_mat
)

# Empirical-Bayes moderated t-tests per contrast.
propeller_full_list <- lapply(colnames(contr_mat), function(cn) {
  res <- speckle::propeller.ttest(
    prop.list = prop_obj,
    design    = design_mat,
    contrasts = contr_mat[, cn],
    robust    = TRUE,
    trend     = FALSE,
    sort      = FALSE
  )
  data.frame(
    contrast = cn,
    cluster  = rownames(res),
    t_stat   = res$Tstatistic,
    p_value  = res$P.Value,
    fdr      = res$FDR,
    stringsAsFactors = FALSE
  )
})

# Tidy propeller results restricted to the three erythroid stages.
propeller_df <- dplyr::bind_rows(lapply(propeller_full_list, function(res) {
  parts <- strsplit(unique(res$contrast), "_vs_")[[1]]
  data.frame(
    ery_stage     = res$cluster,
    group1        = parts[1],
    group2        = parts[2],
    p_propeller   = res$p_value,
    fdr_propeller = res$fdr,
    stringsAsFactors = FALSE
  )
})) %>%
  dplyr::filter(ery_stage %in% names(marker_panels))
```

```{r per_replicate_fractions}
# Per-replicate stage fractions (of total splenocytes) for the bars/dots.
total_per_sample <- so@meta.data %>%
  dplyr::count(orig.ident, genotype, name = "n_total_sample")

stage_frac_df <- so_ery@meta.data %>%
  dplyr::count(orig.ident, genotype, ery_stage, name = "n_stage") %>%
  tidyr::complete(tidyr::nesting(orig.ident, genotype),
                  ery_stage = names(marker_panels),
                  fill = list(n_stage = 0)) %>%
  dplyr::left_join(total_per_sample, by = "orig.ident") %>%
  dplyr::mutate(
    genotype       = factor(dplyr::coalesce(as.character(genotype.x),
                                            as.character(genotype.y)),
                            levels = geno_levels),
    ery_stage      = factor(ery_stage, levels = names(marker_panels)),
    stage_fraction = n_stage / n_total_sample
  ) %>%
  dplyr::select(orig.ident, genotype, ery_stage, stage_fraction)
```

---

# Publication figure (per-stage fractions)

The bar plot below is the manuscript figure: it shows the direction and
magnitude of the genotype-level fractions per maturation stage, with
propeller p-values reported in brackets.

```{r build_figure, fig.width=7.2, fig.height=3.4}
nice_stage <- c(Erythroid_Progenitor = "Progenitor",
                Erythroblast         = "Erythroblast",
                Late_Erythroblast    = "Late erythroblast")

# Per-genotype summary: mean +/- SEM across replicates.
stage_frac_summary <- stage_frac_df %>%
  dplyr::group_by(genotype, ery_stage) %>%
  dplyr::summarise(mean_frac = mean(stage_fraction),
                   sem       = sd(stage_fraction) / sqrt(dplyr::n()),
                   .groups   = "drop")

# Significance brackets, one per pairwise contrast, stacked per facet.
fmt_p <- function(p) ifelse(p < 1e-3, sprintf("p = %.1e", p), sprintf("p = %.3f", p))

y_stage_max <- stage_frac_df %>%
  dplyr::group_by(ery_stage) %>%
  dplyr::summarise(y_max = max(stage_fraction), .groups = "drop")

bracket_df <- propeller_df %>%
  dplyr::left_join(y_stage_max, by = "ery_stage") %>%
  dplyr::mutate(
    ery_stage = factor(ery_stage, levels = names(marker_panels)),
    label     = fmt_p(p_propeller),
    x1        = match(group1, geno_levels),
    x2        = match(group2, geno_levels),
    x_mid     = (x1 + x2) / 2
  ) %>%
  dplyr::group_by(ery_stage) %>%
  dplyr::arrange(abs(x2 - x1), x1, .by_group = TRUE) %>%
  dplyr::mutate(
    step  = y_max * 0.26,
    tick  = y_max * 0.045,
    y_bar = y_max + dplyr::row_number() * step,
    y_txt = y_bar + tick * 1.3
  ) %>%
  dplyr::ungroup()

p_prop_fig <- ggplot(stage_frac_summary,
                     aes(x = genotype, y = mean_frac, fill = genotype)) +
  geom_col(colour = "black", linewidth = 0.3, width = 0.62) +
  geom_errorbar(aes(ymin = pmax(mean_frac - sem, 0), ymax = mean_frac + sem),
                width = 0.2, linewidth = 0.4) +
  geom_jitter(data = stage_frac_df,
              aes(x = genotype, y = stage_fraction),
              width = 0.12, height = 0, size = 1.6, shape = 21,
              fill = "white", colour = "black", stroke = 0.4,
              alpha = 0.9, inherit.aes = FALSE) +
  # Significance brackets (propeller p-values).
  geom_segment(data = bracket_df, aes(x = x1, xend = x2, y = y_bar, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = bracket_df, aes(x = x1, xend = x1, y = y_bar - tick, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = bracket_df, aes(x = x2, xend = x2, y = y_bar - tick, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_text(data = bracket_df, aes(x = x_mid, y = y_txt, label = label),
            size = 2.6, inherit.aes = FALSE) +
  facet_wrap(~ ery_stage, scales = "free_y",
             labeller = labeller(ery_stage = nice_stage)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1),
                     expand = expansion(mult = c(0.02, 0.45))) +
  scale_fill_manual(values = geno_cols, guide = "none") +
  labs(x = NULL, y = "Fraction of all captured splenocytes") +
  theme_classic(base_size = 11) +
  theme(
    axis.text       = element_text(colour = "black"),
    axis.title.y    = element_text(margin = margin(r = 6)),
    strip.background = element_blank(),
    strip.text      = element_text(face = "bold", size = 11),
    panel.spacing   = unit(1, "lines")
  )

p_prop_fig
```

```{r export_figure}
# Export publication-grade copies: vector (PDF, SVG) + 300 dpi raster (PNG, TIFF).
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
fig_base <- file.path(FIG_DIR, "erythroid_propeller_composition")

ggsave(paste0(fig_base, ".pdf"), p_prop_fig,
       width = 7.2, height = 3.4, device = cairo_pdf)
ggsave(paste0(fig_base, ".png"), p_prop_fig,
       width = 7.2, height = 3.4, dpi = 300)
ggsave(paste0(fig_base, ".tiff"), p_prop_fig,
       width = 7.2, height = 3.4, dpi = 300, compression = "lzw")
try(ggsave(paste0(fig_base, ".svg"), p_prop_fig,
           width = 7.2, height = 3.4), silent = TRUE)

message("Figure written to: ", normalizePath(FIG_DIR))
```

**Figure (primary). Splenic erythroid composition across CASP8
genotypes.** Per-stage fraction of total captured splenocytes in WT, KO and
CS spleens (all *Ripk3^-/-^*), faceted by erythroid maturation stage. Bars,
mean across biological replicates (n = 2-4 mice per genotype); error bars,
+/- s.e.m. (standard error of the mean); dots, individual animals. Brackets give propeller p-values
(`speckle::propeller.ttest`; arcsine-sqrt, empirical-Bayes moderation;
Phipson *et al.* 2022). All three erythroid stages are increased in CS
relative to WT and KO, consistent with stress-driven splenic
erythropoiesis in *Casp8^CS/CS^/Ripk3^-/-^* mice.

---

# Diagnostic figure — propeller QQ

For each pairwise contrast (CS vs WT, CS vs KO, WT vs KO), the propeller
moderated t-statistic of each splenic cluster is plotted against its rank
quantile. The three erythroid stages are highlighted in red; the pooled
non-erythroid compartment ("Other") appears as the compositional complement
in grey. This is an internal diagnostic of the per-stage t-statistics and is
not used in the manuscript.

```{r build_qq_figure, fig.width=9, fig.height=3.6}
qq_df <- dplyr::bind_rows(propeller_full_list) %>%
  dplyr::group_by(contrast) %>%
  dplyr::mutate(
    n        = dplyr::n(),
    rank     = rank(t_stat),
    expected = qnorm((rank - 0.5) / n),
    is_ery   = cluster %in% names(marker_panels),
    cluster_label = dplyr::recode(cluster,
                                  Erythroid_Progenitor = "Progenitor",
                                  Erythroblast         = "Erythroblast",
                                  Late_Erythroblast    = "Late erythroblast")
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(contrast = gsub("_vs_", " vs ", contrast))

p_qq_fig <- ggplot(qq_df, aes(expected, t_stat)) +
  # Reference lines: null diagonal + zero lines for sign orientation.
  geom_hline(yintercept = 0, colour = "grey85", linewidth = 0.3) +
  geom_vline(xintercept = 0, colour = "grey85", linewidth = 0.3) +
  geom_abline(slope = 1, intercept = 0, colour = "grey55",
              linetype = "dashed", linewidth = 0.4) +
  # Null cloud: quiet, small, semi-transparent.
  geom_point(data = dplyr::filter(qq_df, !is_ery),
             colour = "grey55", fill = "grey75", shape = 21,
             size = 1.6, alpha = 0.6, stroke = 0.2) +
  # Erythroid points: foreground, red fill with thin black border.
  geom_point(data = dplyr::filter(qq_df, is_ery),
             fill = "#D73027", colour = "black", shape = 21,
             size = 3, stroke = 0.35) +
  ggrepel::geom_text_repel(
    data  = dplyr::filter(qq_df, is_ery),
    aes(label = cluster_label),
    colour       = "#A50026",
    fontface     = "bold",
    size         = 2.9,
    box.padding  = 0.6,
    point.padding = 0.3,
    min.segment.length = 0,
    segment.size       = 0.3,
    segment.colour     = "grey45",
    seed = 42, max.overlaps = Inf
  ) +
  # Dummy aes only to drive the legend (two-row swatch).
  geom_point(data = data.frame(
               expected = NA_real_, t_stat = NA_real_,
               grp = c("Erythroid stage", "Other splenic cluster")),
             aes(fill = grp, shape = grp),
             colour = "black", size = 2.5, stroke = 0.3, na.rm = TRUE) +
  scale_fill_manual(
    name   = NULL,
    values = c("Erythroid stage" = "#D73027",
               "Other splenic cluster" = "grey75")
  ) +
  scale_shape_manual(
    name   = NULL,
    values = c("Erythroid stage" = 21, "Other splenic cluster" = 21)
  ) +
  facet_wrap(~ contrast, nrow = 1) +
  labs(
    x        = "Theoretical quantile (standard normal)",
    y        = "Propeller t-statistic",
    subtitle = "Positive t \u2192 higher proportion in the first genotype of the contrast"
  ) +
  theme_classic(base_size = 11) +
  theme(
    axis.text        = element_text(colour = "black"),
    plot.subtitle    = element_text(size = 9, colour = "grey35",
                                    margin = margin(b = 6)),
    strip.background = element_rect(fill = "grey92", colour = NA),
    strip.text       = element_text(face = "bold", size = 11,
                                    margin = margin(t = 3, b = 3)),
    legend.position  = "bottom",
    legend.key.size  = unit(0.5, "lines"),
    panel.spacing    = unit(1.1, "lines")
  ) +
  guides(
    fill  = guide_legend(override.aes = list(size = 2.8, stroke = 0.3)),
    shape = guide_legend(override.aes = list(size = 2.8, stroke = 0.3))
  )

p_qq_fig
```

```{r export_qq_figure}
# Export copies of the diagnostic QQ figure.
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
qq_base <- file.path(FIG_DIR, "erythroid_propeller_qq")

ggsave(paste0(qq_base, ".pdf"), p_qq_fig,
       width = 9, height = 3.6, device = cairo_pdf)
ggsave(paste0(qq_base, ".png"), p_qq_fig,
       width = 9, height = 3.6, dpi = 300)
ggsave(paste0(qq_base, ".tiff"), p_qq_fig,
       width = 9, height = 3.6, dpi = 300, compression = "lzw")
try(ggsave(paste0(qq_base, ".svg"), p_qq_fig,
           width = 9, height = 3.6), silent = TRUE)

message("QQ figure written to: ", normalizePath(FIG_DIR))
```

**Figure (diagnostic). Per-stage propeller t-statistics across pairwise
genotype contrasts.** Empirical-Bayes moderated t-statistics
(`speckle::propeller.ttest`; arcsine-sqrt transform, robust variance;
Phipson *et al.* 2022) for the three erythroid maturation stages (red) and
the pooled non-erythroid splenocyte compartment ("Other"; grey, shown as
compositional complement), plotted against standard-normal rank quantiles
for each pairwise contrast. Positive t -> higher proportion in the first
genotype; dashed line, identity. In CS-containing panels, all three
erythroid stages show positive t-statistics, consistent with expanded
splenic erythropoiesis in *Casp8^CS/CS^/Ripk3^-/-^* mice. Internal
diagnostic; effect sizes and p-values are reported in the publication bar
plot and statistics table.

---

# Source data (per-replicate fractions)

The table below contains the per-replicate stage fractions underlying the
figure. Use the **Excel** button to download.

```{r source_data_table}
source_data <- stage_frac_df %>%
  dplyr::transmute(
    sample   = orig.ident,
    genotype = genotype,
    stage    = dplyr::recode(as.character(ery_stage), !!!nice_stage),
    fraction = round(stage_fraction, 5)
  ) %>%
  dplyr::arrange(stage, genotype, sample)

DT::datatable(
  source_data,
  caption   = "Per-replicate stage fractions of total splenocytes.",
  rownames  = FALSE,
  extensions = "Buttons",
  options   = list(
    dom      = "Bfrtip",
    buttons  = list(list(extend = "excel",
                         filename = "erythroid_propeller_source_data",
                         title = NULL)),
    pageLength = 30
  )
)
```

---

# Statistics table

```{r stats_table}
stats_out <- propeller_df %>%
  dplyr::transmute(
    stage             = dplyr::recode(ery_stage, !!!nice_stage),
    contrast          = paste(group1, "vs", group2),
    `p (propeller)`   = signif(p_propeller, 3),
    `FDR (propeller)` = signif(fdr_propeller, 3)
  ) %>%
  dplyr::arrange(stage, contrast)

DT::datatable(
  stats_out,
  caption    = "Pairwise propeller tests (arcsine-sqrt + empirical-Bayes variance moderation, two-sided contrasts). FDR = Benjamini–Hochberg across stages within each contrast.",
  rownames   = FALSE,
  extensions = "Buttons",
  options    = list(
    dom      = "Bfrtip",
    buttons  = list(list(extend = "excel",
                         filename = "erythroid_propeller_statistics",
                         title = NULL)),
    pageLength = 15
  )
)
```

```{r export_tables}
# Also write Excel workbook with both sheets for file-based delivery.
wb <- openxlsx::createWorkbook()
openxlsx::addWorksheet(wb, "Source Data")
openxlsx::writeDataTable(wb, "Source Data", source_data)
openxlsx::addWorksheet(wb, "Propeller Statistics")
openxlsx::writeDataTable(wb, "Propeller Statistics", stats_out)
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
xlsx_path <- file.path(FIG_DIR, "erythroid_propeller_tables.xlsx")
openxlsx::saveWorkbook(wb, xlsx_path, overwrite = TRUE)
message("Excel workbook written to: ", xlsx_path)
```

---

# Reproducibility

```{r session_info}
sessionInfo()
```

# References

- Phipson B, Sim CB, Porrello ER, Hewitt AW, Powell J, Oshlack A. *propeller:
  testing for differences in cell type proportions in single cell data.*
  **Bioinformatics** 38(20):4720–4726, 2022.
- Ritchie ME *et al.* *limma powers differential expression analyses for
  RNA-sequencing and microarray studies.* **Nucleic Acids Research** 43(7):e47,
  2015.
- Hao Y *et al.* *Dictionary learning for integrative, multimodal and scalable
  single-cell analysis.* **Nature Biotechnology** 42:293–304, 2024 (Seurat v5).
