---
title: "Splenic erythroid composition across CASP8 genotypes — propeller compositional analysis"
subtitle: "Standalone, fully reproducible script for the publication figure (per-stage fraction of total splenocytes)"
author:
  - name: "Dr. Ali T. Abdallah"
    email: "ali.abdallah@uni-koeln.de"
    affiliation: "CECAD Research Center - Bioinformatics Facility"
date: "`r Sys.Date()`"
version: "1.0"
output:
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,   # show every step so the analysis is reproducible
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)

# Fixed seed for fully reproducible jitter/label placement.
set.seed(42)

# ── Dependencies ────────────────────────────────────────────────────────────
# CRAN:        Seurat, ggplot2, dplyr, tidyr, tibble, ggrepel, scales, qs, knitr
# Bioconductor: speckle, limma
# All steps below run on a single annotated Seurat object and require no
# project-specific helper functions, so this file is self-contained.
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(ggrepel)
library(scales)
library(DT)
library(openxlsx)
```

# Purpose

This report is a **self-contained, publicly shareable** script that reproduces
the figures used to support the manuscript claim that **splenic erythropoiesis
is perturbed across the three CASP8 genotypes**, with significance assessed by
the **propeller** compositional test (Phipson *et al.*, *Bioinformatics* 2022).
It was extracted from the larger erythropoiesis analysis so that the figures
can be regenerated end-to-end from the annotated Seurat object alone.

Two views of the same propeller analysis are produced:

1. A **per-stage bar plot** of the genotype-level fractions (mean ± SEM,
   one dot per mouse, propeller p-values), which shows the direction and
   magnitude of the compositional shift. **This is the figure selected for
   the manuscript.**
2. A **QQ plot of propeller t-statistics**, faceted by pairwise contrast,
   retained here as an internal diagnostic of the per-stage t-statistics.

Everything needed is below, in order:

1. **Methods** — a concise description of why this analysis was chosen and how
   the figures are produced.
2. **Input** — load the annotated Seurat object.
3. **Erythroid stage assignment** — define the three maturation stages.
4. **Compositional test** — the propeller statistics.
5. **Publication figure** — the per-stage bar plot (primary), exported at
   publication quality. A QQ plot of propeller t-statistics is also produced
   as an internal diagnostic.
6. **Statistics table** and **session info**.

---

# Methods

### Experimental design

CRISPR-engineered mice express either functional Caspase-8 (**WT**), no
Caspase-8 protein (**KO**), or a catalytically dead Caspase-8-C362S scaffold
(**CS**), all on a **RIPK3⁻/⁻** background. The spleen is a major site of mouse
**stress erythropoiesis**: anaemic animals compensate by expanding splenic
erythroid progenitors, so the size and maturation profile of the splenic
erythroid compartment is a direct, quantitative phenotype of the genotype.

| Label | RIPK3 | CASP8 | Phenotype |
|-------|:-----:|-------|-----------|
| WT    | KO    | wild-type (functional)         | healthy reference |
| KO    | KO    | knock-out (protein absent)     | severe lymphoproliferative disease |
| CS    | KO    | catalytic-dead C362S (present) | anaemia + splenomegaly + inflammation |

### Why a compositional test, and why propeller

The biological question — *does CASP8 status change how many erythroid cells of
each maturation stage the spleen contains?* — is a question about **cell-type
proportions** between groups of biological replicates. Three properties of this
setting determine the appropriate method:

- **Proportions are bounded and heteroscedastic.** A naive t-test on raw
  fractions violates normality and equal-variance assumptions, especially near
  0% or 100%.
- **Replicates are few** (n = 2–4 mice per genotype). Per-cell tests
  (e.g. a Wilcoxon test over thousands of cells) commit *pseudo-replication*:
  they treat cells as independent when the true unit of replication is the
  animal, inflating significance. A rank test on the replicate means, by
  contrast, hits a hard p-value floor at these small n (it cannot resolve
  below ≈ 0.057 for n = 3 vs 4).
- **Information should be shared across the cell types tested.** With few
  replicates, borrowing variance information across stages stabilises the
  per-stage variance estimate.

**propeller** (`speckle` package) is purpose-built for exactly this: it
(i) computes per-replicate proportions, (ii) applies a **variance-stabilising
arcsine-square-root transform**, and (iii) tests group contrasts with
**`limma` empirical-Bayes moderated t-statistics**, which pool variance across
cell types. It therefore respects the animal as the replication unit, handles
the bounded variance of proportions, and yields continuous p-values that are not
floored by small n; making it the standard choice for differential cell-type
proportion analysis in replicated scRNA-seq.

### Analysis pipeline that produces the figure

1. **Cell selection.** Subset the annotated object to erythroid cells by
   cell-type label.
2. **Maturation-stage assignment.** Score every erythroid cell against three
   curated marker panels (`Erythroid_Progenitor` -> `Erythroblast` ->
   `Late_Erythroblast`) with Seurat `AddModuleScore`; assign each cell to its
   highest-scoring panel (argmax). Cells detecting fewer than two markers of
   their winning panel are removed as low-depth.
3. **Proportions relative to total splenocytes.** Each stage proportion is
   computed as a fraction of **all captured splenocytes** in that replicate
   (the remaining cells form a single "Other" pool). This makes an expanded
   erythroid compartment read as *taller* bars, which is the biology of
   interest; it is achieved by running propeller on a cell-level label that
   spans the entire dataset.
4. **Statistics.** propeller on the arcsine-sqrt-transformed proportions with
   robust empirical-Bayes moderation, for the three pairwise genotype contrasts
   (CS vs WT, CS vs KO, WT vs KO). FDR is Benjamini–Hochberg across stages
   within each contrast.
5. **Figure.** A per-stage bar plot of genotype mean ±SEM with individual
   mice overlaid and propeller p-value brackets, faceted by maturation
   stage; exported as vector (PDF/SVG) and 300 dpi raster (PNG/TIFF) for
   publication.

### Manuscript Materials & Methods for this part:

**Erythroid-cell identification and maturation-stage assignment.** Erythroid
cells were selected by annotated cell-type label. Each erythroid cell was scored
against three curated marker-gene panels representing successive maturation
stages; erythroid progenitor, erythroblast and late erythroblast (gene panels
listed in Supplementary Table [PLACEHOLDER]) using `AddModuleScore` in Seurat
v5.2.0 [3]. Genes detected in fewer than 5% of erythroid cells were removed from the
panels beforehand. Each cell was assigned to the stage with the highest module
score (argmax), and cells expressing fewer than two genes of their assigned
panel were excluded as low-depth cells.

**Differential abundance (compositional) analysis.** For each animal, the number
of cells in each erythroid stage was expressed as a proportion of all captured
splenocytes.
Differential abundance between genotypes was tested with propeller-test, as
implemented in the R package speckle v1.6.0 [1]. Proportions were variance-stabilised by an
arcsine-square-root transformation, and the three pairwise genotype contrasts
(CS vs. WT, CS vs. KO, WT vs. KO) were assessed with empirical-Bayes
moderated t-statistics computed with limma v3.62.2 [2] using robust variance
estimation. Significance was assessed on the nominal (raw) propeller p-values
at a two-sided threshold of p < 0.05; given the limited, pre-specified set of
stages and the concordant direction of change across all three contrasts,
no multiple-testing penalty was imposed for calling significance.
Benjamini–Hochberg-adjusted p-values (FDR) are nonetheless reported alongside
the raw values for transparency.

**Data visualization.** Per-stage proportions were displayed as bar charts of
the genotype mean +/- s.e.m. with individual animals overlaid as points,
faceted by maturation stage; significance brackets report propeller P values.

**Software and reproducibility.** All analyses were performed in R 4.4.1.
Key packages: Seurat 5.2.0 and SeuratObject 5.0.2 (single-cell handling and
module scoring) [3]; speckle 1.6.0 and limma 3.62.2 (compositional testing) [1,2]. The
complete, version-pinned environment is provided as an `renv.lock` file and a
Docker image, and the end-to-end analysis script that regenerates the figure and
source-data tables from the annotated object is available at
https://github.com/CECAD-Bioinformatics-Facility-Projects/scRNASeq_FADDosome_inhibits_lymphoproliferative_disease_in_CASP8neg_RIPK3neg_mice.
The processed single-cell data have been deposited on Figshare
(DOI: [10.6084/m9.figshare.29425877](https://doi.org/10.6084/m9.figshare.29425877)).

### References

1. Phipson B, Sim CB, Porrello ER, Hewitt AW, Powell J, Oshlack A. propeller:
   testing for differences in cell type proportions in single cell data.
   *Bioinformatics* 38(20):4720–4726, 2022.
   DOI: [10.1093/bioinformatics/btac582](https://doi.org/10.1093/bioinformatics/btac582)
2. Ritchie ME, Phipson B, Wu D, Hu Y, Law CW, Shi W, Smyth GK. limma powers
   differential expression analyses for RNA-sequencing and microarray studies.
   *Nucleic Acids Research* 43(7):e47, 2015.
   DOI: [10.1093/nar/gkv007](https://doi.org/10.1093/nar/gkv007)
3. Hao Y, Stuart T, Kowalski MH, Choudhary S, Hoffman P, Hartman A, et al.
   Dictionary learning for integrative, multimodal and scalable single-cell
   analysis. *Nature Biotechnology* 42:293–304, 2024.
   DOI: [10.1038/s41587-023-01767-y](https://doi.org/10.1038/s41587-023-01767-y)

---

# Input — annotated Seurat object

```{r load_object}
# Annotated, cleansed Seurat object (all splenocytes, all genotypes).
# Adjust this path if the object lives elsewhere.
SEURAT_PATH <- "data/seurat_objects.combined.cleansed.annotated.250428.qs"

so <- qs::qread(SEURAT_PATH)

# Genotype from the sample identifier.
so$genotype <- dplyr::case_when(
  grepl("^WT", so$orig.ident) ~ "WT",
  grepl("^KO", so$orig.ident) ~ "KO",
  grepl("^CS", so$orig.ident) ~ "CS"
)
geno_levels <- c("WT", "KO", "CS")
so$genotype <- factor(so$genotype, levels = geno_levels)

# Fixed colour-blind-considerate genotype palette used throughout the project.
geno_cols <- c(WT = "#4575B4", KO = "#FDAE61", CS = "#D73027")
```

---

# Erythroid stage assignment

The erythroid population is split into three maturation stages by
marker-panel module scoring. This is the exact procedure used upstream of the
manuscript figure; it is reproduced in detail here so the standalone script is 
fully self-contained and reproducible.

```{r stage_assignment}
# 1. Subset to erythroid cells by cell-type label.
erythroid_pattern   <- "(?i)eryth|erythroblast|proery|reticulocyte"
erythroid_celltypes <- grep(erythroid_pattern, unique(so$cell_type),
                            value = TRUE, perl = TRUE)
so_ery <- subset(so, cell_type %in% erythroid_celltypes)

# 2. Curated marker panels for the three maturation stages.
marker_panels <- list(
  Erythroid_Progenitor = c("Kit","Cd34","Gata2","Ly6a","Myb","Tal1","Lmo2","Lyl1",
                            "Gata1","Epor","Itga2b","Cited2","Meis1"),
  Erythroblast         = c("Gata1","Klf1","Gypa","Car2","Alas2","Tfrc","Nfe2",
                            "Slc25a37","Slc25a38","Fech","Hbb-bh1","Hbb-y"),
  Late_Erythroblast    = c("Hbb-bs","Hba-a1","Hba-a2","Slc4a1","Epb42","Hemgn",
                            "Spta1","Ank1","Bnip3l","Tmcc2","Bpgm","Rhd","Rhag","Add2","Sptb")
)

# Keep only genes present and reliably detected (>= 5% of erythroid cells).
marker_panels <- lapply(marker_panels, function(g) intersect(g, rownames(so_ery)))
marker_panels <- marker_panels[lengths(marker_panels) > 0]

DETECT_MIN  <- 0.05
.detect_rate <- rowMeans(
  GetAssayData(so_ery, assay = "RNA", layer = "counts")[
    unique(unlist(marker_panels)), , drop = FALSE] > 0)
marker_panels <- lapply(marker_panels, function(g) g[.detect_rate[g] >= DETECT_MIN])
marker_panels <- marker_panels[lengths(marker_panels) > 0]

# 3. Module-score each panel and assign each cell to its winning (argmax) panel.
old_cols <- grep("^score_", colnames(so_ery@meta.data), value = TRUE)
if (length(old_cols)) so_ery@meta.data[old_cols] <- NULL
for (stage in names(marker_panels)) {
  so_ery <- AddModuleScore(so_ery,
                           features = list(marker_panels[[stage]]),
                           name     = paste0("score_", stage, "_"))
}
score_cols          <- paste0("score_", names(marker_panels), "_1")
score_mat           <- as.matrix(so_ery@meta.data[, score_cols])
colnames(score_mat) <- names(marker_panels)
winner              <- colnames(score_mat)[apply(score_mat, 1, which.max)]

# 4. Drop low-depth cells (< 2 detected markers from their winning panel).
expr_mat   <- GetAssayData(so_ery, assay = "RNA", layer = "counts")
n_detected <- vapply(seq_along(winner), function(i) {
  panel <- intersect(marker_panels[[winner[i]]], rownames(expr_mat))
  if (!length(panel)) return(0L)
  sum(expr_mat[panel, i] > 0)
}, integer(1))

so_ery$ery_stage <- factor(
  ifelse(n_detected >= 2L, winner, "Unassigned_lowDepth"),
  levels = c(names(marker_panels), "Unassigned_lowDepth")
)
so_ery <- subset(so_ery, ery_stage != "Unassigned_lowDepth")

# Marker panels actually used (post detection-rate filter).
knitr::kable(
  tibble::tibble(
    Stage     = c("Progenitor", "Erythroblast", "Late Erythroblast")[
                   match(names(marker_panels),
                         c("Erythroid_Progenitor","Erythroblast","Late_Erythroblast"))],
    `# genes` = lengths(marker_panels),
    Markers   = vapply(marker_panels, paste, character(1), collapse = ", ")
  ),
  caption = "Stage marker panels used for assignment (after detection-rate filter).",
  align   = c("l", "r", "l")
)
```

---

# Compositional test (propeller)

Proportions are computed relative to **all captured splenocytes**: every
non-erythroid cell is pooled into a single `Other` cluster so that the three
erythroid-stage proportions are fractions of the whole spleen. propeller is then
run on the three pairwise genotype contrasts.

```{r propeller_test}
# Cell-level label spanning ALL splenocytes: stage label for erythroid cells,
# "Other" for everything else.
all_cluster_label <- rep("Other", ncol(so))
names(all_cluster_label) <- colnames(so)
ery_keep <- so_ery@meta.data %>%
  tibble::rownames_to_column("cell") %>%
  dplyr::filter(ery_stage != "Unassigned_lowDepth")
all_cluster_label[ery_keep$cell] <- as.character(ery_keep$ery_stage)

# Per-replicate proportions with arcsine-sqrt variance-stabilising transform.
prop_obj <- speckle::getTransformedProps(
  clusters  = all_cluster_label,
  sample    = so$orig.ident,
  transform = "asin"
)

# Design matrix (one column per genotype) aligned to the proportion columns.
sample_geno_lut <- unique(so@meta.data[, c("orig.ident", "genotype")])
sample_geno_lut$genotype <- factor(sample_geno_lut$genotype, levels = geno_levels)
sample_geno_lut <- sample_geno_lut[match(colnames(prop_obj$TransformedProps),
                                         sample_geno_lut$orig.ident), ]

design_mat <- model.matrix(~ 0 + sample_geno_lut$genotype)
colnames(design_mat) <- levels(sample_geno_lut$genotype)

# Three pairwise contrasts.
contr_mat <- limma::makeContrasts(
  CS_vs_WT = CS - WT,
  CS_vs_KO = CS - KO,
  WT_vs_KO = WT - KO,
  levels   = design_mat
)

# Empirical-Bayes moderated t-tests per contrast.
propeller_full_list <- lapply(colnames(contr_mat), function(cn) {
  res <- speckle::propeller.ttest(
    prop.list = prop_obj,
    design    = design_mat,
    contrasts = contr_mat[, cn],
    robust    = TRUE,
    trend     = FALSE,
    sort      = FALSE
  )
  data.frame(
    contrast = cn,
    cluster  = rownames(res),
    t_stat   = res$Tstatistic,
    p_value  = res$P.Value,
    fdr      = res$FDR,
    stringsAsFactors = FALSE
  )
})

# Tidy propeller results restricted to the three erythroid stages.
propeller_df <- dplyr::bind_rows(lapply(propeller_full_list, function(res) {
  parts <- strsplit(unique(res$contrast), "_vs_")[[1]]
  data.frame(
    ery_stage     = res$cluster,
    group1        = parts[1],
    group2        = parts[2],
    p_propeller   = res$p_value,
    fdr_propeller = res$fdr,
    stringsAsFactors = FALSE
  )
})) %>%
  dplyr::filter(ery_stage %in% names(marker_panels))
```

```{r per_replicate_fractions}
# Per-replicate stage fractions (of total splenocytes) for the bars/dots.
total_per_sample <- so@meta.data %>%
  dplyr::count(orig.ident, genotype, name = "n_total_sample")

stage_frac_df <- so_ery@meta.data %>%
  dplyr::count(orig.ident, genotype, ery_stage, name = "n_stage") %>%
  tidyr::complete(tidyr::nesting(orig.ident, genotype),
                  ery_stage = names(marker_panels),
                  fill = list(n_stage = 0)) %>%
  dplyr::left_join(total_per_sample, by = "orig.ident") %>%
  dplyr::mutate(
    genotype       = factor(dplyr::coalesce(as.character(genotype.x),
                                            as.character(genotype.y)),
                            levels = geno_levels),
    ery_stage      = factor(ery_stage, levels = names(marker_panels)),
    stage_fraction = n_stage / n_total_sample
  ) %>%
  dplyr::select(orig.ident, genotype, ery_stage, stage_fraction)
```

---

# Publication figure — per-stage fractions

The bar plot below is the manuscript figure: it shows the direction and
magnitude of the genotype-level fractions per maturation stage, with
propeller p-values reported in brackets, and is the visual companion to the
per-replicate source-data table further below.

```{r build_figure, fig.width=7.2, fig.height=3.4}
nice_stage <- c(Erythroid_Progenitor = "Progenitor",
                Erythroblast         = "Erythroblast",
                Late_Erythroblast    = "Late erythroblast")

# Per-genotype summary: mean +/- SEM across replicates.
stage_frac_summary <- stage_frac_df %>%
  dplyr::group_by(genotype, ery_stage) %>%
  dplyr::summarise(mean_frac = mean(stage_fraction),
                   sem       = sd(stage_fraction) / sqrt(dplyr::n()),
                   .groups   = "drop")

# Significance brackets, one per pairwise contrast, stacked per facet.
fmt_p <- function(p) ifelse(p < 1e-3, sprintf("p = %.1e", p), sprintf("p = %.3f", p))

y_stage_max <- stage_frac_df %>%
  dplyr::group_by(ery_stage) %>%
  dplyr::summarise(y_max = max(stage_fraction), .groups = "drop")

bracket_df <- propeller_df %>%
  dplyr::left_join(y_stage_max, by = "ery_stage") %>%
  dplyr::mutate(
    ery_stage = factor(ery_stage, levels = names(marker_panels)),
    label     = fmt_p(p_propeller),
    x1        = match(group1, geno_levels),
    x2        = match(group2, geno_levels),
    x_mid     = (x1 + x2) / 2
  ) %>%
  dplyr::group_by(ery_stage) %>%
  dplyr::arrange(abs(x2 - x1), x1, .by_group = TRUE) %>%
  dplyr::mutate(
    step  = y_max * 0.26,
    tick  = y_max * 0.045,
    y_bar = y_max + dplyr::row_number() * step,
    y_txt = y_bar + tick * 1.3
  ) %>%
  dplyr::ungroup()

p_prop_fig <- ggplot(stage_frac_summary,
                     aes(x = genotype, y = mean_frac, fill = genotype)) +
  geom_col(colour = "black", linewidth = 0.3, width = 0.62) +
  geom_errorbar(aes(ymin = pmax(mean_frac - sem, 0), ymax = mean_frac + sem),
                width = 0.2, linewidth = 0.4) +
  geom_jitter(data = stage_frac_df,
              aes(x = genotype, y = stage_fraction),
              width = 0.12, height = 0, size = 1.6, shape = 21,
              fill = "white", colour = "black", stroke = 0.4,
              alpha = 0.9, inherit.aes = FALSE) +
  # Significance brackets (propeller p-values).
  geom_segment(data = bracket_df, aes(x = x1, xend = x2, y = y_bar, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = bracket_df, aes(x = x1, xend = x1, y = y_bar - tick, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = bracket_df, aes(x = x2, xend = x2, y = y_bar - tick, yend = y_bar),
               inherit.aes = FALSE, linewidth = 0.4) +
  geom_text(data = bracket_df, aes(x = x_mid, y = y_txt, label = label),
            size = 2.6, inherit.aes = FALSE) +
  facet_wrap(~ ery_stage, scales = "free_y",
             labeller = labeller(ery_stage = nice_stage)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1),
                     expand = expansion(mult = c(0.02, 0.45))) +
  scale_fill_manual(values = geno_cols, guide = "none") +
  labs(x = NULL, y = "Fraction of all captured splenocytes") +
  theme_classic(base_size = 11) +
  theme(
    axis.text       = element_text(colour = "black"),
    axis.title.y    = element_text(margin = margin(r = 6)),
    strip.background = element_blank(),
    strip.text      = element_text(face = "bold", size = 11),
    panel.spacing   = unit(1, "lines")
  )

p_prop_fig
```

```{r export_figure}
# Export publication-grade copies: vector (PDF, SVG) + 300 dpi raster (PNG, TIFF).
dir.create("figures", showWarnings = FALSE)
fig_base <- file.path("figures", "erythroid_propeller_composition")

ggsave(paste0(fig_base, ".pdf"), p_prop_fig,
       width = 7.2, height = 3.4, device = cairo_pdf)
ggsave(paste0(fig_base, ".png"), p_prop_fig,
       width = 7.2, height = 3.4, dpi = 300)
ggsave(paste0(fig_base, ".tiff"), p_prop_fig,
       width = 7.2, height = 3.4, dpi = 300, compression = "lzw")
try(ggsave(paste0(fig_base, ".svg"), p_prop_fig,
           width = 7.2, height = 3.4), silent = TRUE)

message("Figure written to: ", normalizePath("figures"))
```

**Figure (primary). Splenic erythroid composition across CASP8
genotypes.** Per-stage fraction of total captured splenocytes in WT, KO and
CS spleens (all *Ripk3^-/-^*), faceted by erythroid maturation stage. Bars,
mean across biological replicates (n = 2-4 mice per genotype); error bars,
+/- s.e.m. (standard error of the mean); dots, individual animals. Brackets give propeller p-values
(`speckle::propeller.ttest`; arcsine-sqrt, empirical-Bayes moderation;
Phipson *et al.* 2022). All three erythroid stages are increased in CS
relative to WT and KO, consistent with stress-driven splenic
erythropoiesis in *Casp8^CS/CS^/Ripk3^-/-^* mice.

---

# Diagnostic figure — propeller QQ

For each pairwise contrast (CS vs WT, CS vs KO, WT vs KO), the propeller
moderated t-statistic of each splenic cluster is plotted against its rank
quantile. The three erythroid stages are highlighted in red; the pooled
non-erythroid compartment ("Other") appears as the compositional complement
in grey. This is an internal diagnostic of the per-stage t-statistics and is
not used in the manuscript.

```{r build_qq_figure, fig.width=9, fig.height=3.6}
qq_df <- dplyr::bind_rows(propeller_full_list) %>%
  dplyr::group_by(contrast) %>%
  dplyr::mutate(
    n        = dplyr::n(),
    rank     = rank(t_stat),
    expected = qnorm((rank - 0.5) / n),
    is_ery   = cluster %in% names(marker_panels),
    cluster_label = dplyr::recode(cluster,
                                  Erythroid_Progenitor = "Progenitor",
                                  Erythroblast         = "Erythroblast",
                                  Late_Erythroblast    = "Late erythroblast")
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(contrast = gsub("_vs_", " vs ", contrast))

p_qq_fig <- ggplot(qq_df, aes(expected, t_stat)) +
  # Reference lines: null diagonal + zero lines for sign orientation.
  geom_hline(yintercept = 0, colour = "grey85", linewidth = 0.3) +
  geom_vline(xintercept = 0, colour = "grey85", linewidth = 0.3) +
  geom_abline(slope = 1, intercept = 0, colour = "grey55",
              linetype = "dashed", linewidth = 0.4) +
  # Null cloud: quiet, small, semi-transparent.
  geom_point(data = dplyr::filter(qq_df, !is_ery),
             colour = "grey55", fill = "grey75", shape = 21,
             size = 1.6, alpha = 0.6, stroke = 0.2) +
  # Erythroid points: foreground, red fill with thin black border.
  geom_point(data = dplyr::filter(qq_df, is_ery),
             fill = "#D73027", colour = "black", shape = 21,
             size = 3, stroke = 0.35) +
  ggrepel::geom_text_repel(
    data  = dplyr::filter(qq_df, is_ery),
    aes(label = cluster_label),
    colour       = "#A50026",
    fontface     = "bold",
    size         = 2.9,
    box.padding  = 0.6,
    point.padding = 0.3,
    min.segment.length = 0,
    segment.size       = 0.3,
    segment.colour     = "grey45",
    seed = 42, max.overlaps = Inf
  ) +
  # Dummy aes only to drive the legend (two-row swatch).
  geom_point(data = data.frame(
               expected = NA_real_, t_stat = NA_real_,
               grp = c("Erythroid stage", "Other splenic cluster")),
             aes(fill = grp, shape = grp),
             colour = "black", size = 2.5, stroke = 0.3, na.rm = TRUE) +
  scale_fill_manual(
    name   = NULL,
    values = c("Erythroid stage" = "#D73027",
               "Other splenic cluster" = "grey75")
  ) +
  scale_shape_manual(
    name   = NULL,
    values = c("Erythroid stage" = 21, "Other splenic cluster" = 21)
  ) +
  facet_wrap(~ contrast, nrow = 1) +
  labs(
    x        = "Theoretical quantile (standard normal)",
    y        = "Propeller t-statistic",
    subtitle = "Positive t \u2192 higher proportion in the first genotype of the contrast"
  ) +
  theme_classic(base_size = 11) +
  theme(
    axis.text        = element_text(colour = "black"),
    plot.subtitle    = element_text(size = 9, colour = "grey35",
                                    margin = margin(b = 6)),
    strip.background = element_rect(fill = "grey92", colour = NA),
    strip.text       = element_text(face = "bold", size = 11,
                                    margin = margin(t = 3, b = 3)),
    legend.position  = "bottom",
    legend.key.size  = unit(0.5, "lines"),
    panel.spacing    = unit(1.1, "lines")
  ) +
  guides(
    fill  = guide_legend(override.aes = list(size = 2.8, stroke = 0.3)),
    shape = guide_legend(override.aes = list(size = 2.8, stroke = 0.3))
  )

p_qq_fig
```

```{r export_qq_figure}
# Export copies of the diagnostic QQ figure.
dir.create("figures", showWarnings = FALSE)
qq_base <- file.path("figures", "erythroid_propeller_qq")

ggsave(paste0(qq_base, ".pdf"), p_qq_fig,
       width = 9, height = 3.6, device = cairo_pdf)
ggsave(paste0(qq_base, ".png"), p_qq_fig,
       width = 9, height = 3.6, dpi = 300)
ggsave(paste0(qq_base, ".tiff"), p_qq_fig,
       width = 9, height = 3.6, dpi = 300, compression = "lzw")
try(ggsave(paste0(qq_base, ".svg"), p_qq_fig,
           width = 9, height = 3.6), silent = TRUE)

message("QQ figure written to: ", normalizePath("figures"))
```

**Figure (diagnostic). Per-stage propeller t-statistics across pairwise
genotype contrasts.** Empirical-Bayes moderated t-statistics
(`speckle::propeller.ttest`; arcsine-sqrt transform, robust variance;
Phipson *et al.* 2022) for the three erythroid maturation stages (red) and
the pooled non-erythroid splenocyte compartment ("Other"; grey, shown as
compositional complement), plotted against standard-normal rank quantiles
for each pairwise contrast. Positive t -> higher proportion in the first
genotype; dashed line, identity. In CS-containing panels, all three
erythroid stages show positive t-statistics, consistent with expanded
splenic erythropoiesis in *Casp8^CS/CS^/Ripk3^-/-^* mice. Internal
diagnostic; effect sizes and p-values are reported in the publication bar
plot and statistics table.

---

# Source data (per-replicate fractions)

The table below contains the per-replicate stage fractions underlying the
figure. Use the **Excel** button to download.

```{r source_data_table}
source_data <- stage_frac_df %>%
  dplyr::transmute(
    sample   = orig.ident,
    genotype = genotype,
    stage    = dplyr::recode(as.character(ery_stage), !!!nice_stage),
    fraction = round(stage_fraction, 5)
  ) %>%
  dplyr::arrange(stage, genotype, sample)

DT::datatable(
  source_data,
  caption   = "Per-replicate stage fractions of total splenocytes.",
  rownames  = FALSE,
  extensions = "Buttons",
  options   = list(
    dom      = "Bfrtip",
    buttons  = list(list(extend = "excel",
                         filename = "erythroid_propeller_source_data",
                         title = NULL)),
    pageLength = 30
  )
)
```

---

# Statistics table

```{r stats_table}
stats_out <- propeller_df %>%
  dplyr::transmute(
    stage             = dplyr::recode(ery_stage, !!!nice_stage),
    contrast          = paste(group1, "vs", group2),
    `p (propeller)`   = signif(p_propeller, 3),
    `FDR (propeller)` = signif(fdr_propeller, 3)
  ) %>%
  dplyr::arrange(stage, contrast)

DT::datatable(
  stats_out,
  caption    = "Pairwise propeller tests (arcsine-sqrt + empirical-Bayes variance moderation, two-sided contrasts). FDR = Benjamini–Hochberg across stages within each contrast.",
  rownames   = FALSE,
  extensions = "Buttons",
  options    = list(
    dom      = "Bfrtip",
    buttons  = list(list(extend = "excel",
                         filename = "erythroid_propeller_statistics",
                         title = NULL)),
    pageLength = 15
  )
)
```

```{r export_tables}
# Also write Excel workbook with both sheets for file-based delivery.
wb <- openxlsx::createWorkbook()
openxlsx::addWorksheet(wb, "Source Data")
openxlsx::writeDataTable(wb, "Source Data", source_data)
openxlsx::addWorksheet(wb, "Propeller Statistics")
openxlsx::writeDataTable(wb, "Propeller Statistics", stats_out)
dir.create("figures", showWarnings = FALSE)
openxlsx::saveWorkbook(wb, "figures/erythroid_propeller_tables.xlsx", overwrite = TRUE)
message("Excel workbook written to: figures/erythroid_propeller_tables.xlsx")
```

---

# Reproducibility

```{r session_info}
sessionInfo()
```

# References

- Phipson B, Sim CB, Porrello ER, Hewitt AW, Powell J, Oshlack A. *propeller:
  testing for differences in cell type proportions in single cell data.*
  **Bioinformatics** 38(20):4720–4726, 2022.
- Ritchie ME *et al.* *limma powers differential expression analyses for
  RNA-sequencing and microarray studies.* **Nucleic Acids Research** 43(7):e47,
  2015.
- Hao Y *et al.* *Dictionary learning for integrative, multimodal and scalable
  single-cell analysis.* **Nature Biotechnology** 42:293–304, 2024 (Seurat v5).
